---
title: "Untitled"
output: html_document
---
# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
# setting up environment
rm(list=ls())
# setwd("~/Yale/2022 Fall/Tobin RA")
# need to make a new wd now
# loading packages
library(readxl)
library(tidyverse)
library(lubridate)
library(Hmisc)
library(heplots)
library(ggExtra)
library(caret)
library(class)

# loading data
load("alldata.RData")
# load("allyears.RData")
# dfList <- list(df_2015, df_2016, df_2017, df_2018, df_2019, df_2020)

```


# Checking Counts

## Checking Gender Counts
```{r}
### sanity check for gender counts ###
genderdist <- function(dat){
  hist1 <- ggplot(dat, aes(x=GENDER)) +
    geom_bar(stat="count", width=0.7, fill="steelblue") +
    theme(aspect.ratio=0.7)
  return(hist1)
}

# genderdist(df_2015)
# genderdist(df_2016)
# genderdist(df_2017)
# genderdist(df_2018)
# genderdist(df_2019)
# genderdist(df_2020)

for (i in 1:length(2015:2020)){
  iter <- filter(alldata, year(alldata$UPDATE_DATE) == c(2015:2020)[i])
  print(genderdist(iter))
}

for (i in 1:length(2015:2020)){
  iter <- filter(alldata, year(alldata$UPDATE_DATE) == c(2015:2020)[i])
  print(table(as.factor(iter$GENDER)))
}


table(alldata$GENDER)
```

There are a fairly even number of responses from both males and females across the years. It seems the proportion of "Other" resposnes has been slightly increasing. This could be due to better clarity for the option, shifts in social norms, or maybe better anti-discriminatory policy.

There are many NAs, we should discuss what to do with these later.

## Checking unique responses
```{r}
### checking if there are duplicates within each year
########## deprecated code based on having a dataframe for each year, 
########## too lazy to change it haha
# length(unique(df_2015$PID))-nrow(df_2015)
# length(unique(df_2016$PID))-nrow(df_2016)
# length(unique(df_2017$PID))-nrow(df_2017)
# length(unique(df_2018$PID))-nrow(df_2018)
# length(unique(df_2019$PID))-nrow(df_2019)
# length(unique(df_2020$PID))-nrow(df_2020)
# no duplicates within year
```

There are no duplicates within each year.

# Checking dates
```{r}
str(alldata$UPDATE_DATE) 
# already in a date format!!

# no NAs across all years (I deleted code for this)

# organizing dates by month
# df_2015$newdates <- floor_date(df_2015$UPDATE_DATE, unit = "month")
# df_2016$newdates <- floor_date(df_2016$UPDATE_DATE, unit = "month")
# df_2017$newdates <- floor_date(df_2017$UPDATE_DATE, unit = "month")
# df_2018$newdates <- floor_date(df_2018$UPDATE_DATE, unit = "month")
# df_2019$newdates <- floor_date(df_2019$UPDATE_DATE, unit = "month")
# df_2020$newdates <- floor_date(df_2020$UPDATE_DATE, unit = "month")


# seeing how dates are dispersed, checking for weird ones
date_hist <- function(x){
  hist1 <- ggplot(data = x, aes(x = floor_date(UPDATE_DATE, unit = "month"))) +
    geom_bar(stat = "count", fill="steelblue") 
  return(hist1)
}



for (i in 1:length(2015:2020)){
  iter <- filter(alldata, year(alldata$UPDATE_DATE) == c(2015:2020)[i])
  iterhist <- date_hist(iter) + 
    xlab(as.character(c(2015:2020)[i]))
  print(iterhist)
}
```

The first few years had some weird spreads of dates, I'll check these further individually.

```{r}
#table(df_2015$newdates)
```

The dataframe for 2016 had survey response dates goinog all the way to december of 2017. I'm guessing we should remove these?

```{r}
#table(df_2016$newdates)
```

The same is true for 2016.

```{r}
#table(df_2017$newdates)
```

2017 had 2 extraneous dates from 2016. Weird.

All in all, things seem to be mostly in order. 


# Checking Industry

```{r}
# dfList %>%
#     lapply(., function(x) sum(is.na(x$Industry))) %>%
#     unlist() /
#   (dfList %>%
#     lapply(., function(x) nrow(x)) %>%
#     unlist())

sum(is.na(alldata$Industry)) / nrow(alldata)
# about 12% missing data
```

About 12% of data for Industry is missing.

Let's see some breakdowns of these industries
```{r}
indhist <- function(x){
  hist1 <- ggplot(data=x, aes(x=Industry)) +
    geom_bar(stat="count", width=0.7, fill="steelblue")
  return(hist1)
}

sort(table(alldata$Industry))
sum(is.na(alldata$Industry)) / nrow(alldata)
```

This is kind of messy, I'm not sure if we want to really get into the weeds with this yet. 

# Checking yrs exp

```{r}
exphist <- function(x){
  hist1 <- ggplot(data=x, aes(x=YRS_EXP))+
    geom_density(stat="count", alpha=0.6, fill="steelblue")
}

for (i in 1:length(2015:2020)){
  iter <- filter(alldata, year(alldata$UPDATE_DATE) == c(2015:2020)[i])
  print(exphist(iter))
}
```

Overall the spreads of work experience are right skewed, which makes sense.

# Checking hourly_rate
```{r}
describe(alldata$HOURLY_RATE)
quantile(alldata$HOURLY_RATE, .99, na.rm=TRUE)
```
95th, 99th percentile makes sense, there are a few outliers that may be from very heavily salaried employees.

# Checking TCC spread
```{r}
salaryhist <- function(x){
  hist1 <- ggplot(data=x, aes(x=TCC)) +
    geom_density(fill = 'steelblue', alpha=0.75) +
    xlim(c(0,401622))
}

#sample stats
describe(alldata$TCC)
quantile(alldata$TCC, 0.99, na.rm = TRUE)

for (i in 1:length(2015:2020)){
  iter <- filter(alldata, year(alldata$UPDATE_DATE) == c(2015:2020)[i])
  print(salaryhist(iter))
}
```

There's a ton of outliers! We should make a cutoff for this. I've set it to 401,622, the 99th percentile income in 2020.

# Gender Comparisons

## TCC by gender
```{r}
# overlaid histogram for salary vs gender
salaryhist2 <- function(x){
  x %>%
    filter(!is.na(GENDER)) %>%
  ggplot() +
    geom_density(aes(x = TCC, fill = GENDER), #TCC has much less NA than salary
                 alpha=0.25) + 
    theme(legend.position = c(0.1, .73), # adding legend
          plot.title = element_text(vjust = 0.25),
          aspect.ratio = 0.75) + # moving title
    xlim(c(0,200000)) +
    theme_minimal() +
    ylim(0, 3.5 * 10 ** (-5)) +
    scale_fill_manual(values = c("#FF0000", "#169B45", "#0F31D8", "#5B6CC1")) 
}

for (i in 1:length(2015:2020)){
  iter <- filter(alldata, year(alldata$UPDATE_DATE) == c(2015:2020)[i])
  print(salaryhist2(iter))
}

```

```{r}

# a function for comparing TCC for missing data to non-missing
salaryhist3 <- function(x){
  x %>%
    filter(!is.na(GENDER)) %>%
  ggplot() +
    geom_density(aes(x = TCC, fill = (GENDER=="Missing")), #TCC has much less NA than salary
                 alpha=0.25) + 
    theme(legend.position = c(0.1, .73), # adding legend
          plot.title = element_text(vjust = 0.25),
          aspect.ratio = 0.75) + # moving title
    xlim(c(0,200000)) +
    theme_minimal() +
    ylim(0, 3.5 * 10 ** (-5)) +
    scale_fill_manual(values = c("#FF0000", "#169B45")) 
}

for(i in 1:length(2015:2020)){
  alldata %>% 
    filter(year(alldata$UPDATE_DATE) == c(2015:2020)[i]) %>% 
    salaryhist3() %>% 
    print()
}
```


```{r}
# Gender data vs missing
ggplot(alldata) +
    geom_density(aes(x = TCC, fill = GENDER=="Missing"),alpha=0.5)


```


```{r}
# generating summary stats for TCC
# female
alldata %>% filter(GENDER=="Female") %>% select(TCC) %>%
  Hmisc::describe(.)
alldata %>% filter(GENDER=="Female") %>% select(TCC) %>% unlist() %>% sd()

# male
alldata %>% filter(GENDER=="Male") %>% select(TCC) %>%
  Hmisc::describe(.)
alldata %>% filter(GENDER=="Male") %>% select(TCC) %>% unlist() %>% sd()

# other
alldata %>% filter(GENDER=="Other") %>% select(TCC) %>%
  Hmisc::describe(.)
alldata %>% filter(GENDER=="Other") %>% select(TCC) %>% unlist() %>% sd()
```

## Salary by gender
```{r}
mean(alldata$SALARY[alldata$GENDER == "Female"], na.rm = TRUE)
mean(alldata$SALARY[alldata$GENDER == "Male"], na.rm = TRUE)
for (i in 1:length(2015:2020)){
  tempfem <- alldata %>%
    filter(GENDER=="Female") %>%
    filter(year(UPDATE_DATE) == c(2015:2020)[i]) %>%
    select(SALARY) %>%
    colMeans(na.rm = TRUE)
  tempmale <- alldata %>%
    filter(GENDER=="Male") %>%
    filter(year(UPDATE_DATE) == c(2015:2020)[i]) %>%
    select(SALARY) %>%
    colMeans(na.rm = TRUE)
  print(tempfem - tempmale)
}


```



## Industry by gender
```{r}
# industry breakdown across all year
table(alldata$Industry, alldata$GENDER)

#breakdown across multiple years
for (i in 1:length(2015:2020)){
  iter <- filter(alldata, year(alldata$UPDATE_DATE) == c(2015:2020)[i])
  print(round(prop.table(table(iter$Industry, iter$GENDER),2),5))
}
sort(unique(alldata$Industry))

# add horiz barplot, use naics codes only
```

```{r}
# Top industries by gender
head(sort(table(alldata$Industry[alldata$GENDER=="Female"]),decreasing = TRUE))
head(sort(table(alldata$Industry[alldata$GENDER=="Male"]),decreasing = TRUE))
head(sort(table(alldata$Industry[alldata$GENDER=="Other"]),decreasing = TRUE))
head(sort(table(alldata$Industry[alldata$GENDER!="Missing"]),decreasing = TRUE))
head(sort(table(alldata$Industry[alldata$GENDER=="Missing"]),decreasing = TRUE))
```


## Management by gender
```{r}
table(alldata$MGMT_ROLE, alldata$GENDER)

alldata %>% filter(MGMT_ROLE == "Management Role") %>%
  ggplot(aes(x=as.factor(year(UPDATE_DATE)),y=MGMT_ROLE)) +
    geom_bar(stat="identity",aes(fill=GENDER)) + 
    ggtitle("management role by gender")


alldata %>% filter(MGMT_ROLE == "Management Role") %>%
  ggplot(aes(x=as.factor(year(UPDATE_DATE)),y=MGMT_ROLE)) +
    geom_col(aes(fill=GENDER), position = "fill") + 
    ggtitle("management role by gender") #+
    #ylim(c(0,1.2))

# another way for a single year
barplot(prop.table(table(alldata$MGMT_ROLE, alldata$GENDER), margin=2))


# ggplot(alldata, aes(x=as.factor(year(UPDATE_DATE)),y=MGMT_ROLE)) +
#   geom_bar(stat = "identity",aes(fill=GENDER))
```
 

## Educ by gender
```{r}
table(alldata$EDU_LVL, alldata$GENDER)
```

Pretty even spread! Females tend to get more associate's, health doctorate, 
master's, and certificates than men. This may be due to more women responding
to the survey, though. Let's maybe try to find a better measure of this (e.g.
by proportion).

```{r}
round(prop.table(table(alldata$EDU_LVL, alldata$GENDER), margin = 2),5)
```

These seem more balanced, but the above differences seem to still persist.

Let's observe any changes over time

```{r}
# for (i in 1:length(2015:2020)){
#   iter <- filter(alldata, year(alldata$UPDATE_DATE) == c(2015:2020)[i])
#   print(round(prop.table(table(iter$EDU_LVL, iter$GENDER),margin=1),6))
# }
```

```{r}
# summary stats for education level

# proportion tables
round(prop.table(table(alldata$highest,alldata$GENDER), margin = 2),4)
#comparing missing with complete gender data
round(prop.table(table(alldata$highest[alldata$GENDER %in% c("Male", "Female", "Other")])),4)
round(prop.table(table(alldata$highest[alldata$GENDER == "Missing"])),4)

# NA counts
sum(is.na(alldata$EDU_LVL[alldata$GENDER=="Female"]))
sum(is.na(alldata$EDU_LVL[alldata$GENDER=="Male"]))
sum(is.na(alldata$EDU_LVL[alldata$GENDER=="Other"]))
```


## Yrs exp by gender
```{r}
# generating summary stats for yrs_exp
# female
alldata %>% filter(GENDER=="Female") %>% select(YRS_EXP) %>%
  Hmisc::describe(.)
sd(alldata$YRS_EXP[alldata$GENDER=="Female"], na.rm = TRUE)

# male
alldata %>% filter(GENDER=="Male") %>% select(YRS_EXP) %>%
  Hmisc::describe(.)
sd(alldata$YRS_EXP[alldata$GENDER=="Male"], na.rm = TRUE)

# Other
alldata %>% filter(GENDER=="Other") %>% select(YRS_EXP) %>%
  Hmisc::describe(.)
sd(alldata$YRS_EXP[alldata$GENDER=="Other"], na.rm = TRUE)
```


## Age by gender
```{r}
agehist <- function(x){
  x %>%
    filter(!is.na(GENDER) & GENDER %in% c("Female", "Male")) %>% # remove NAs
  ggplot() +
    geom_density(aes(x = AGE, fill = GENDER), 
                 alpha=0.25) + 
    theme(legend.position = c(0.1, .73), # adding legend
          plot.title = element_text(vjust = 0.25),
          aspect.ratio = 0.75) + # moving title
    theme_minimal() +
    scale_fill_manual(values = c("#FF0000", "#169B45", "#0F31D8")) 
}

for (i in 1:length(2015:2020)){
  iter <- filter(alldata, year(alldata$UPDATE_DATE) == c(2015:2020)[i])
  print(agehist(iter))
}


agehist(alldata)
```
 
 Ages are pretty left skewed, which we can take into account when considering
 external effects of broader trends.


```{r}
# generating summary stats for yrs_exp
# female
alldata %>% filter(GENDER=="Female") %>% select(AGE) %>%
  Hmisc::describe(.)
sd(alldata$AGE[alldata$GENDER=="Female"], na.rm = TRUE)

# male
alldata %>% filter(GENDER=="Male") %>% select(AGE) %>%
  Hmisc::describe(.)
sd(alldata$AGE[alldata$GENDER=="Male"], na.rm = TRUE)

# Other
alldata %>% filter(GENDER=="Other") %>% select(AGE) %>%
  Hmisc::describe(.)
sd(alldata$AGE[alldata$GENDER=="Other"], na.rm = TRUE)
```


## Salary hist question by gender
```{r}
# female
prop.table(table(alldata$SALARY_HISTORY[alldata$GENDER=="Female"]))

# male
prop.table(table(alldata$SALARY_HISTORY[alldata$GENDER=="Male"]))

# Other
prop.table(table(alldata$SALARY_HISTORY[alldata$GENDER=="Other"]))
# 2:54
# 1:26

# missing data per year
for (i in 1:length(2015:2020)){
    iter <- filter(alldata, year(alldata$UPDATE_DATE) == c(2015:2020)[i])
    print(sum(is.na(iter$SALARY_HISTORY)))
}
# total obs per year
for (i in 1:length(2015:2020)){
    iter <- filter(alldata, year(alldata$UPDATE_DATE) == c(2015:2020)[i])
    print(length(iter$SALARY_HISTORY))
}
```

## Company size
```{r}
# female
alldata %>% filter(GENDER=="Female") %>% select(COMPANY_SIZE) %>%
  Hmisc::describe(.)
sd(alldata$COMPANY_SIZE[alldata$GENDER=="Female"], na.rm = TRUE)
# male
alldata %>% filter(GENDER=="Male") %>% select(COMPANY_SIZE) %>%
  Hmisc::describe(.)
sd(alldata$COMPANY_SIZE[alldata$GENDER=="Male"], na.rm = TRUE)
# other
alldata %>% filter(GENDER=="Other") %>% select(COMPANY_SIZE) %>%
  Hmisc::describe(.)
sd(alldata$COMPANY_SIZE[alldata$GENDER=="Other"], na.rm = TRUE)
```

## Hourly Workweek by Gender
```{r}
alldata %>% filter(GENDER=="Female") %>% select(HOURLY_WORKWEEK) %>%
  Hmisc::describe(.)
sd(alldata$HOURLY_WORKWEEK[alldata$GENDER=="Female"], na.rm = TRUE)
# male
alldata %>% filter(GENDER=="Male") %>% select(HOURLY_WORKWEEK) %>%
  Hmisc::describe(.)
sd(alldata$HOURLY_WORKWEEK[alldata$GENDER=="Male"], na.rm = TRUE)
# other
alldata %>% filter(GENDER=="Other") %>% select(HOURLY_WORKWEEK) %>%
  Hmisc::describe(.)
sd(alldata$HOURLY_WORKWEEK[alldata$GENDER=="Other"], na.rm = TRUE)
```


```{r}
# chi-square quantile plot function, credit to JDRS
#A function to make chi-square quantile plots 
#to test for multivariate normality of data or residuals
CSQPlot<-function(vars,label="Chi-Square Quantile Plot"){
  #usually, vars is xxx$residuals or data from one group and label is for plot
  x<-cov(scale(vars),use="pairwise.complete.obs")
  squares<-sort(diag(as.matrix(scale(vars))%*%solve(x)%*%as.matrix(t(scale(vars)))))
  quantiles<-quantile(squares)
  hspr<-quantiles[4]-quantiles[2]
  cumprob<-c(1:length(vars[,1]))/length(vars[,1])-1/(2*length(vars[,1]))
  degf<-dim(x)[1]
  quants<-qchisq(cumprob,df=degf)
  gval<-(quants**(-1+degf/2))/(exp(quants/2)*gamma(degf/2)*(sqrt(2)**degf))
  scale<-hspr / (qchisq(.75,degf)-qchisq(.25,degf))
  se<-(scale/gval)*sqrt(cumprob*(1-cumprob)/length(squares))
  lower<-quants-2*se
  upper<-quants+2*se
  
  plot(quants,squares,col='red',pch=19,cex=1.2,xlab="Chi-Square Quantiles",
       ylab="Squared MH Distance",main=paste("Chi-Square Quantiles for",label),ylim=range(upper,lower, squares) , xlim=range(c(0,quants)))
  lines(c(0,100),c(0,100),col=1)
  lines(quants,upper,col="blue",lty=2,lwd=2)
  lines(quants,lower,col="blue",lty=2,lwd=2)
  legend("topleft",c("Data","95% Conf Limits"),lty=c(0,2),col=c("red","blue"),lwd=c(2,2),
         pch=c(19,NA))
}
#An example of usage of this function
#CSQPlot(danielaaov$residuals,label="Daniela MANOVA Residuals")
```

Note: tests for multivariate normality must have shorter tables, should I
just sample them?

```{r}

#CSQPlot(num2015) #error: not enough ram
#heplots::cqplot(num2015)
# library(QuantPsyc)
# mult.norm(num2015)$mult.test
# CSQPlot(sample_n(num2015,500))

#cor(num2015$HOURLY_RATE, num2015$SALARY, use="complete.obs")
# interesting error - is it either salary or wage reported?
```


# States and Cities
```{r}
sort(prop.table(table(alldata$LOCATION_STATE)), TRUE)
```
A good mix of SHB and non-SHB states

A possible cool way to plot state data, probably not necessary (maybe a side proj)
https://urban-institute.medium.com/how-to-create-state-and-county-maps-easily-in-r-577d29300bb2


```{r}
head(sort(prop.table(table(alldata$LOCATION_CITY)),TRUE),15)
```

Here's the 15 most prominent cities in our dataset. 


# SHQ Comparisons

## TCC by SHQ

```{r}
# generating summary stats for TCC
# not missing
alldata %>% filter(!is.na(SALARY_HISTORY)) %>% select(TCC) %>%
  Hmisc::describe(.)
alldata %>% filter(!is.na(SALARY_HISTORY)) %>% select(TCC) %>% unlist() %>% sd()

# missing
alldata %>% filter(is.na(SALARY_HISTORY)) %>% select(TCC) %>%
  Hmisc::describe(.)
alldata %>% filter(is.na(SALARY_HISTORY)) %>% select(TCC) %>% unlist() %>% sd()

```


## Industry by SHQ
```{r}
# industry breakdown across all year
table(alldata$Industry, is.na(alldata$SALARY_HISTORY))

```

```{r}
# Top industries by SHQ
head(sort(table(alldata$Industry[!is.na(alldata$SALARY_HISTORY)]),decreasing = TRUE))
head(sort(table(alldata$Industry[is.na(alldata$SALARY_HISTORY)]),decreasing = TRUE))
```


## Management by SHQ
```{r}
table(alldata$MGMT_ROLE, is.na(alldata$SALARY_HISTORY))

alldata %>% filter(MGMT_ROLE == "Management Role") %>%
  ggplot(aes(x=as.factor(year(UPDATE_DATE)),y=MGMT_ROLE)) +
    geom_bar(stat="identity",aes(fill=GENDER)) + 
    ggtitle("management role by gender")


alldata %>% filter(MGMT_ROLE == "Management Role") %>%
  ggplot(aes(x=as.factor(year(UPDATE_DATE)),y=MGMT_ROLE)) +
    geom_col(aes(fill=GENDER), position = "fill") + 
    ggtitle("management role by gender") #+
    #ylim(c(0,1.2))

# another way for a single year
barplot(prop.table(table(alldata$MGMT_ROLE, alldata$GENDER), margin=2))


# ggplot(alldata, aes(x=as.factor(year(UPDATE_DATE)),y=MGMT_ROLE)) +
#   geom_bar(stat = "identity",aes(fill=GENDER))
```
 

## Educ by SHQ

```{r}
# summary stats for education level
# proportion tables
round(prop.table(table(alldata$highest, is.na(alldata$SALARY_HISTORY)), margin = 2),4)
sum(is.na(alldata$highest[!is.na(alldata$SALARY_HISTORY)]))
sum(is.na(alldata$highest[is.na(alldata$SALARY_HISTORY)]))
```


## Yrs exp by SHQ
```{r}
# generating summary stats for yrs_exp
# not missing
alldata %>% filter(!is.na(SALARY_HISTORY)) %>% select(YRS_EXP) %>%
  Hmisc::describe(.)
sd(alldata$YRS_EXP[!is.na(alldata$SALARY_HISTORY)], na.rm = TRUE)

# missing
alldata %>% filter(is.na(SALARY_HISTORY)) %>% select(YRS_EXP) %>%
  Hmisc::describe(.)
sd(alldata$YRS_EXP[is.na(alldata$SALARY_HISTORY)], na.rm = TRUE)
```


## Age by SHQ

```{r}
# generating summary stats for age
# not missing
alldata %>% filter(!is.na(SALARY_HISTORY)) %>% select(AGE) %>%
  Hmisc::describe(.)
sd(alldata$AGE[!is.na(alldata$SALARY_HISTORY)], na.rm = TRUE)

# missing
alldata %>% filter(is.na(SALARY_HISTORY)) %>% select(AGE) %>%
  Hmisc::describe(.)
sd(alldata$AGE[is.na(alldata$SALARY_HISTORY)], na.rm = TRUE)
```


## Salary hist question
```{r}
# not missing
prop.table(table(alldata$SALARY_HISTORY[!is.na(alldata$SALARY_HISTORY)]))

```

## Company size
```{r}
# female
alldata %>% filter(GENDER=="Female") %>% select(COMPANY_SIZE) %>%
  Hmisc::describe(.)
sd(alldata$COMPANY_SIZE[alldata$GENDER=="Female"], na.rm = TRUE)
# male
alldata %>% filter(GENDER=="Male") %>% select(COMPANY_SIZE) %>%
  Hmisc::describe(.)
sd(alldata$COMPANY_SIZE[alldata$GENDER=="Male"], na.rm = TRUE)
# other
alldata %>% filter(GENDER=="Other") %>% select(COMPANY_SIZE) %>%
  Hmisc::describe(.)
sd(alldata$COMPANY_SIZE[alldata$GENDER=="Other"], na.rm = TRUE)
```

## Hourly Workweek by Gender
```{r}
alldata %>% filter(GENDER=="Female") %>% select(HOURLY_WORKWEEK) %>%
  Hmisc::describe(.)
sd(alldata$HOURLY_WORKWEEK[alldata$GENDER=="Female"], na.rm = TRUE)
# male
alldata %>% filter(GENDER=="Male") %>% select(HOURLY_WORKWEEK) %>%
  Hmisc::describe(.)
sd(alldata$HOURLY_WORKWEEK[alldata$GENDER=="Male"], na.rm = TRUE)
# other
alldata %>% filter(GENDER=="Other") %>% select(HOURLY_WORKWEEK) %>%
  Hmisc::describe(.)
sd(alldata$HOURLY_WORKWEEK[alldata$GENDER=="Other"], na.rm = TRUE)
```




# a prelim knn for gender

Note: doesn't produce good results

```{r}
# subset data with dependent and indep vars
# remove incomplete data
mldata <- alldata %>% select(TCC, AGE, highest, LOCATION_STATE, GENDER,
                             ONET_BROAD) %>% 
  na.omit()
# subset size for data for computational speed
mldata <-  mldata[sample(nrow(mldata), 5000), ]
# turning factors into numeric for knn function
mldata$highest <- as.numeric(as.factor(mldata$highest))
mldata$LOCATION_STATE <- as.numeric(as.factor(mldata$LOCATION_STATE))
mldata$ONET_BROAD <- as.numeric(as.factor(mldata$ONET_BROAD))

# partitioning data into training and test data
sample <- sample(c(TRUE, FALSE), nrow(mldata), replace=TRUE, prob=c(0.7,0.3))
training_x  <- mldata[sample, -5]
testing_x   <- mldata[!sample, -5]
training_y <- as.factor(mldata$GENDER[sample])
testing_y <- as.factor(mldata$GENDER[!sample])


# knn based on 10 closest neighbors
Mn <- knn(train=training_x,test=testing_x,
             cl=training_y,k=10)

# accuracey for 10 neighbors
mean(testing_y==Mn)

# accuracy for 1 to 20 neighbors
result <- c()
for (i in 1:20) {
   set.seed(1)
   Mn <- knn(train=training_x,test=testing_x,
             cl=training_y,k=i)
   result[i]<- mean(testing_y==Mn)
}
result
```

# prelim multinomial logistic
```{r}
index <- createDataPartition(mldata$GENDER, p = .70, list = FALSE)
train <- mldata[index,]
test <- mldata[-index,]
train$GENDER <- as.factor(train$GENDER)
test$GENDER <- as.factor(test$GENDER)
train$GENDER <- relevel(train$GENDER, ref = "Male")

multinom_model <- nnet::multinom(GENDER ~ ., data = train)

test$predicted <- predict(multinom_model, newdata = test, "class")
tab <- table(test$GENDER, test$predicted)
round((sum(diag(tab))/sum(tab))*100,2)
```



# Instrumental variable

A proposed model using 2 state least-square (2SLS): 

Stage 1: Modeling salary disclosure:

$$ \widehat{Disc}_{n \times 1} = \delta _{0 _ {n \times 1}} + 
\delta _{1} SHB _ {n \times 1} + \gamma _{r\times 1} V _{n \times r}
$$

Where V is a matrix of other control variables


$$ Y=\beta_0 + \beta _1 \widehat{Disc} + \gamma V + \epsilon $$

Where $Y$ is total compensation,  $X$ is a matrix of controls, $\beta_c$ is 
the vector of coefficients for
the controls, $SHB$ is the vector of pre/post SHB, and $\beta_SHB$ is the 
coefficient for SHB. We also have $Gender*SHB*\beta_{inter}$, which accounts
for the interaction of gender and disclosure. 
Here, SHB is a instrument for salary disclosure, as other
unobserved variables may impact both salary history disclosure *and* other
independent variables. 


First let's check for potential interaction effects
```{r}
interaction.plot(x.factor = alldata$GENDER, #x-axis variable
                 trace.factor = alldata$SHB, #variable for lines
                 response = alldata$TCC, #y-axis variable
                 fun = median, #metric to plot
                 ylab = "Total Compensation",
                 xlab = "Gender",
                 col = c("red", "blue"),
                 lty = 1, #line type
                 lwd = 2, #line width
                 trace.label = "SHB?")
```


We can test for the inclusion restriction of instrumental variables via 
first-stage regression, i.e. regressing disclosure on SHB.

```{r}
# focus on (question+disclosure) and (no question+no disclosure)
alldata2 <- filter(alldata, 
                   SALARY_HISTORY %in% c("No, and they did not ask",
                                         "Yes, they asked about my salary history"))
alldata2$SALARY_HISTORY <- as.numeric(as.factor(alldata2$SALARY_HISTORY))-1 #1=NO, 2=YES
stage1 <- lm(SALARY_HISTORY ~ SHB + GENDER + AGE + RACE_SET + highest + LOCATION_STATE + ONET_MINOR, data = alldata2)
summary1 <- stage1 %>% summary()
summary1$coefficients[c(1,2,3,6),]
```

Next we can look at relationships(or lack thereof) between our instrument (SHB) 
and our other controls. We can do this with covariances, plots, and basic linear models.

```{r}
#covariances for numeric, binary controls
cov(alldata$SHB, as.numeric(as.factor(alldata$GENDER)))
cov(alldata$SHB, alldata$AGE)
```


```{r}
# first some plots (gross code sorry)
sample1 <- alldata2[sample(nrow(alldata2),10000),]
sample1 <- sample1[sample1$RACE_SET %in% c("American Indian and Alaska Native",
                                           "Asian",
                                           "Black or African American",
                                           "Hispanic",
                                           "White"),]
# salary history
plot(jitter(SALARY_HISTORY, 0.5)~jitter(SHB, 0.5), data = sample1)
abline(lm(SALARY_HISTORY~SHB, data = sample1)$coefficients[1],
       lm(SALARY_HISTORY~SHB, data = sample1)$coefficients[2])
# gender
plot(jitter(as.numeric(as.factor(GENDER)), 0.5)~jitter(SHB, 0.5), data = sample1,
     ylim=c(0,2.5))
abline(lm(as.numeric(as.factor(GENDER))~SHB, data = alldata2)$coefficients[1],
       lm(as.numeric(as.factor(GENDER))~SHB, data = alldata2)$coefficients[2])
# age
plot(AGE~SHB, data = sample1)
abline(lm(AGE~SHB, data = alldata2)$coefficients[1],
       lm(AGE~SHB, data = alldata2)$coefficients[2])
# race breakdown
sample1 %>%  
  group_by(RACE_SET, SHB) %>%  
  summarise(Count=n()) %>% 
  ggplot(aes(x=RACE_SET, y=Count, fill=as.factor(SHB))) +
  geom_bar(stat='identity', position= "dodge") ### make the sample size 10000


summary(lm(as.numeric(as.vector(SALARY_HISTORY)) ~ SHB, data = alldata2))
summary(lm(as.numeric(as.factor(GENDER))~SHB, data = alldata2))
summary(lm(AGE ~ SHB, data = alldata2))
```



```{r}
# instrumental variable

# subset data and recode salary history question as a binary
alldata2 <- filter(
  alldata,
  SALARY_HISTORY %in% c("No, and they did not ask",
                        "Yes, they asked about my salary history")
)

alldata2$SALARY_HIST_BI <- alldata2$SALARY_HISTORY %>% 
  as.factor() %>%
  as.numeric()
alldata2$SALARY_HIST_BI <- alldata2$SALARY_HIST_BI - 1

# controlled by years experience, race, highest education, state, age, job
# no interactions
library(AER)
iv1 <- ivreg(
  log(SALARY) ~ SALARY_HIST_BI + GENDER + YRS_EXP + AGE + RACE_SET + highest +
    LOCATION_STATE + ONET_MINOR|
    SHB + GENDER + YRS_EXP + AGE + RACE_SET + highest +
    LOCATION_STATE + ONET_MINOR,
   data = alldata2
)

nrow(summary(iv1)$coefficients)
summary(iv1)$coefficients[c(1:7),]



# controlled by years experience, highest education, state, age, job
# no interactions
iv2 <- ivreg(
  log(SALARY) ~ SALARY_HIST_BI + GENDER + YRS_EXP + AGE + highest +
    LOCATION_STATE + ONET_MINOR|
    SHB + GENDER + YRS_EXP + AGE + highest +
    LOCATION_STATE + ONET_MINOR,
   data = alldata2
)

nrow(summary(iv2)$coefficients)
summary(iv2)$coefficients[c(1:7),]



# controlled by years experience, highest education, state, age, job
# with interactions
# different subsetting and factor methods (original)
iv3 <- ivreg(
  log(SALARY) ~ SALARY_HISTORY*GENDER + YRS_EXP + AGE + RACE_SET  + highest +
    LOCATION_STATE + ONET_MINOR|
    SHB*GENDER + GENDER + YRS_EXP + RACE_SET + AGE + highest +
    LOCATION_STATE + ONET_MINOR,
   data = filter(
     alldata, 
     SALARY_HISTORY %in% c("No, and they did not ask",
                           "Yes, they asked about my salary history"))
)

nrow(summary(iv3)$coefficients)
summary(iv3)$coefficients[c(1:7,388:390),]
```

```{r}
stage1 <- lm(as.numeric(as.factor(SALARY_HISTORY)) ~ as.numeric(SHB), data = alldata)
predsalaryhist <- predict.lm(stage1)

iv1 <- lm( 5)
```



# OLS

```{r}
######### don't use this
# ols model, no log adjustment, small control set, no interaction
alldata$ONET_MINOR <- as.factor(alldata$ONET_MINOR)
lm1 <- lm(
  log(TCC) ~ SALARY_HISTORY + GENDER + YRS_EXP + highest + ONET_MINOR,
  # using responses in prof adlers example
  data = filter(
    alldata,
    SALARY_HISTORY %in% c("No, and they did not ask",
                          "Yes, they asked about my salary history")
  ) 
)

summary(lm1)
```

```{r}
# ols model, log salary, no interactions
## controls: age, job, location(state), highest level of education, 
## years of experience
ols1 <- lm(
  log(SALARY) ~ GENDER + SALARY_HISTORY + YRS_EXP + AGE + highest +
    LOCATION_STATE + ONET_MINOR,
  data = alldata
)



summary(ols1)$coefficients[c(1:10),]
```

```{r}
# ols model, log salary, interactions w/ disclosure
## controls: age, job, location(state), highest level of education, 
## years of experience
ols2 <- lm(
  log(SALARY) ~ GENDER*SALARY_HISTORY + YRS_EXP + AGE + highest +
    LOCATION_STATE + ONET_MINOR,
  data = alldata
)

nrow(summary(ols2)$coefficients)
summary(ols2)$coefficients[c(1:10,155:166),]

# subset based on highest present racial groups
ols3 <- alldata %>% filter(RACE_SET %in% c("American Indian and Alaska Native",
                                           "Asian",
                                           "Black or African American",
                                           "Hispanic",
                                           "White")) %>%
  lm(
    log(SALARY) ~ GENDER*SALARY_HISTORY + YRS_EXP + AGE + RACE_SET + highest +
    LOCATION_STATE + ONET_MINOR,
    data = .
  )



summary(ols3)$coefficients[c(1:10,156:168),] # coefficients of interest

```

# DID

```{r}
# year column
alldata$year <- year(alldata$UPDATE_DATE)
table(alldata$year)
table(alldata$SHB)
str(alldata$SHB)
```


```{r}
# DID model, log salary, no interactions
## controls: age, job, location(state), highest level of education, 
## years of experience
did1 <- lm(
  log(SALARY) ~ SHB*GENDER + YRS_EXP + AGE + highest +
    LOCATION_STATE + ONET_MINOR,
  data = alldata
)

nrow(summary(did1)$coefficients)
summary(did1)$coefficients[c(1:8,807),]
```

```{r}
# DID following davis, ouimet, and wang
did2 <- lm(
  log(SALARY) ~ year + SHB*GENDER + YRS_EXP + AGE + highest +
    LOCATION_STATE + ONET_MINOR,
  data = alldata
)
nrow(summary(did2)$coefficients)
summary(did2)$coefficients[c(1:8,156:158),]
```




instrumental var with binary endogenous regressor?

just run ols and instrumental

just get numbers to work with for now

test for instrument relevance

once men or women disclose, what does it do to their pay? (doesn't really depend on motivations for disclosing)

look at tsls documentation in stata



joe peck urban institute
wash inst for equitable growth
brookings
rand corporation




merge in data about local labor market conditions (what did I do over the summer?)
webscraping?

beep boop
